### 2019-02-14 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Okay je réorganise et réécris pas mal pour ce chapitre 4.

@@

% ------------------------------------------------------------------------
\subsection{Motivation for this first IoT model}

Unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.

Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
As end-devices battery life is a key constraint of IoT networks,
this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
%
In this Section, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
Kaufmann12Thompson}.
% Both algorithms have already been applied with success, for Opportunistic Spectrum Access \cite{Jouini10} and recently for multi-users Cognitive Radio problems \cite{Toldov16}.

MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
For example, \cite{Jouini10} presents the OSA setting, and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov16}, and the \UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi17}.
Many recent works show that MAB algorithms work well for real-world radio signal.
However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \emph{i.i.d} (independent and identically distributed).
In recent works like \cite{Bonnefoi17}, several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d}.
As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.

@@

As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance the finite number of available RF channels.
In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station. If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part.
If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.50]{2-Chapters/4-Chapter/CrownCom_17.git/protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

@@

@@

Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :
\begin{subequations}
\label{eq:41:prob}
\begin{align}
\underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
\text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
& D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
\end{align}
\end{subequations}

@@

\begin{equation}\label{eq:41:Dilambda}
D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{\dag} .
\end{equation}
\end{proposition}
\begin{proof}
\begin{itemize}
    \item
    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;K \rrbracket$. For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
        \item
The Lagrange multipliers method can be applied to the optimization problem \eqref{eq:41:optPb}, with a concave objective function $f$, linear equality constraints \eqref{eq:41:eqCstr} and linear inequality constraints \eqref{eq:41:ineqCstr}. The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
    \hfill{}$\square$
    \end{itemize}
\end{proof}

@@

% A well-studied version of this problem is the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be independent and identically distributed (\emph{i.i.d}) under some distribution $\nu_k$, that has a mean $\mu_k$. Several types of reward distributions have been considered, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
% We consider Bernoulli bandit models, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k$.

@@

Considered alone, each dynamic device implements a learning algorithm to play a bandit game, the device is consequently a smart device. In each time slot, if it has to communicate (which happens with probability $p$), then it chooses a channel and it receives a reward $1$ if the transmission is successful, $0$ otherwise.
Each device aims at maximizing the sum of the rewards collected during its communication instants, which shall indeed maximize the fraction of successful transmissions. Besides the modified time scale (rewards are no longer collected at every time step), this looks like a bandit problem.
However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \emph{i.i.d}: they not only depend on the (stationary, \emph{i.i.d}) behavior of the static devices, but also on the behavior of other smart devices, that is not stationary (because of learning).

@@

@@

\paragraph{\emph{Adversarial} bandit algorithms?}
% \todo[inline]{Explain that we also tried adversarial algorithms ($\mathrm{Exp}3$), and it was working but less efficiently than \UCB{} and Thompson Sampling. Maybe add it on some curve?}
Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform worse than both \UCB{} and TS in the same experiments.
%(see Section \ref{sub:41:numericalResults}).
Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary and non-\emph{i.i.d}, as its performance guarantee are true in \emph{any} setting.
But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
%\footnote{We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible. For stochastic algorithms, being ``efficient'' means having a regret bounded as $R_T = \mathcal{O}(\log T)$, but for adversarial algorithms, it means having $R_T = \mathcal{O}(\sqrt{K T})$.}
worse than the one for stochastic ones.
% (in their respective case of application).
More is left on this aspect for our future work.

@@

\begin{figure*}[!t]
    \centering
    \subfloat[10\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/10intelligent.eps}
    \label{fig:41:10intelligent}}
    \hfill
    \subfloat[30\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/30intelligent.eps}
    \label{fig:41:30intelligent}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/50intelligent.eps}
    \label{fig:41:50intelligent}}
    \hfill
    \subfloat[100\% of smart devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/100intelligent.eps}
    \label{fig:41:100intelligent}}
    \caption{Performance of $2$ MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of smart end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure*}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.65]{2-Chapters/4-Chapter/CrownCom_17.git/perf_learning.eps}
    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices.}
    \label{fig:41:perf_learning}
\end{figure}

@@

We include another simulation, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
% And we should include another example with a different scenarios, maybe $(0.3, 0.3, 0.3, 0.015, 0.015, 0.014, 0.014, 0.014, 0.014, 0.014)$ (R\'emi already did the simulations).

\begin{figure*}[!t]
    \centering
    \subfloat[10\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_10.eps}
    \label{fig:41:ch2_10}}
    \hfill
    \subfloat[30\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_30.eps}
    \label{fig:41:ch2_30}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_50.eps}
    \label{fig:41:ch2_50}}
    \hfill
    \subfloat[100\% of intelligent devices]{\includegraphics[scale=0.5]{2-Chapters/4-Chapter/CrownCom_17.git/ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption{Performance of $2$ bandit algorithms, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
    \label{fig:41:figure4appendix}
\end{figure*}

\subsection{Conclusion about this first model}\label{sub:41:conclusion}

In this Section, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the many ``intelligent'' end-devices.

This is both a surprising and a very encouraging result, showing that application of bandit algorithms tailored for a stochastic model is still useful in broader settings.
The fully \emph{decentralized} application of classic stochastic MAB algorithms are almost as efficient as the best possible centralized policy in this setting, after a short learning period, even though the dynamic devices \emph{can not} communicate with each others, and \emph{do not} know the system parameters.
We will investigate this behavior in order to understand it better theoretically.
We will also experiment more with adversarial algorithms, to confirm that they work less efficiently than stochastic bandit algorithms in our non-stochastic setting.

Moreover, for sake of simplicity we supposed that all devices use the same standard. Our future work will consider more realistic interference scenarios and IoT networks, with, \emph{e.g.}, non-slotted time, more than one base station etc.

\paragraph{Possible extensions of this model}
%
This work could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.

Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \emph{e.g.}, one transmission a day.
So additionally to deciding the channel for communication, each device has to also decide when to communicate.
This is another direction of research, that we will investigate in the future.

>>>>

\graphicspath{{2-Chapters/4-Chapter/CrownCom_17.git/}}

@@

% % ------------------------------------------------------------------------
% \subsection{Motivation for this first IoT model}

@@

As illustrated in Figure \ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
In each time slot, devices try to send packets to the unique Base Station, which listens continuously to all channels, following an ALOHA-based communication (no sensing).
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the base station. If only one packet is sent in this part of the slot, the base station can decode it and sends an acknowledgement to the device in the second part.
If two or more devices send an uplink packet in the same slot, the uplink packets collide (\ie, there is a \emph{collision}), and the acknowledgement \emph{Ack} is not transmitted.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.45]{protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fix-duration uplink slot in which the end-devices transmit their packets. If a packet is well received, the base station replies by transmitting an \emph{Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System model - FIXME merge with previous section!}

We consider the system model presented in Figure~\ref{fig:42:system_model1}, where a set of object sends uplink packets to the network gateway, in the $433.5\;\mathrm{MHz}$ ISM band.
The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
%
The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$). In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s).

These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks. This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.

%To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
%
%By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
%the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{system_model1.eps}
    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
    \label{fig:42:system_model1}
\end{figure}

We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
as it is usually done in MAB problems \cite{Thompson33,Robbins52,LaiRobbins85}.

This problem is a special case of the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$. Several types of reward distributions have been considered in the literature, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).

Rewards are binary in our model, and so we consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model. Rewards come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.

The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to the end-users, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
%
This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.

@@

Consequently, the optimal allocation vector $(D_1,\dots,D_{K})$ is the solution of the following real-valued constraint optimization problem :

\begin{equation}
\begin{subequations} \label{eq:41:prob}
    \begin{cases}
    \underset{D_1,\dots,D_{K}}{\arg\max}\; & \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}, \label{eq:41:optPb}\\
    \text{such that}\;\; & \sum_{i=1}^{K} D_i = D, \label{eq:41:eqCstr}\\
    & D_i \geq 0 \qquad \forall i\in\llbracket 1;K\rrbracket . \label{eq:41:ineqCstr}
    \end{cases}
\end{subequations}
\end{equation}

@@

    \begin{equation}\label{eq:41:Dilambda}
        D_i^*(\lambda) = \left(\frac{1}{\log(1-p)}\left[ \mathcal{W}\left(\frac{\lambda e}{(1-p)^{S_i-1}} \right)-1 \right]\right)^{\dag} .
    \end{equation}
\end{proposition}

\begin{proof}
\begin{itemize}
    \item
    In a realistic scenario, we can assume that $D_i\leq \frac{-2}{\ln\left(1-p\right)} \approx \frac{2}{p},\quad \forall i\in\llbracket 1;K \rrbracket$. For such values for $D_i$, the objective function $f: (D_1, \dots, D_{K}) \mapsto \sum_{i=1}^{K} D_i (1 - p)^{S_i + D_i -1}$ is concave as the sum of concave functions
    \footnote{It worth noting that $f$ is neither concave nor quasi-concave on $[0,\infty)^{K}$ \cite{Luenberger68,Yaari77}.}.
    \item
    The Lagrange multipliers method can be applied to the optimization problem \eqref{eq:41:optPb}, with a concave objective function $f$, linear equality constraints \eqref{eq:41:eqCstr} and linear inequality constraints \eqref{eq:41:ineqCstr}. The strong duality condition is satisfied in this case \cite{BoydVanderberghe04}, so finding the saddle points will be enough to find the maximizers.
    \item
    More details are given in Section~\ref{sec:5:proofLagrangian} in the Appendix of this Chapter.
    % \hfill{}$\square$
    \end{itemize}
\end{proof}

@@

% A well-studied version of this problem is the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be independent and identically distributed (\iid) under some distribution $\nu_k$, that has a mean $\mu_k$. Several types of reward distributions have been considered, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).
% We consider Bernoulli bandit models, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k$.

@@

Considered alone, each dynamic device implements a learning algorithm to play a bandit game, the device is consequently a smart device. In each time slot, if it has to communicate (which happens with probability $p$), then it chooses a channel and it receives a reward $1$ if the transmission is successful, $0$ otherwise.
Each device aims at maximizing the sum of the rewards collected during its communication instants, which shall indeed maximize the fraction of successful transmissions. Besides the modified time scale (rewards are no longer collected at every time step), this looks like a bandit problem.
However, it cannot be modeled as a stochastic MAB, as the rewards are clearly \emph{not} \iid: they not only depend on the (stationary, \iid) behavior of the static devices, but also on the behavior of other smart devices, that is not stationary (because of learning).

@@

\paragraph{Two algorithms}

% \paragraph{The \UCB{} Algorithm}\label{sub:42:UCB}

% A naive approach could be to use an empirical mean estimator of the rewards for each channel, and select the channel with highest estimated mean at each time;
% but this greedy approach is known to fail dramatically \cite{LaiRobbins85}. Indeed, with this policy, the selection of arms is highly dependent on the first draws:
% if the first transmission in one channel fails and the first one on other channels succeed, the end-user will \emph{never} use the first channel again, even it is the best one (\emph{i.e.}, the most available, in average).

% Rather than relying on the empirical mean reward, Upper Confidence Bounds algorithms
% instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm,
% which can be viewed as adding a ``bonus'' exploration to the empirical mean.
% % by adding a confidence term to the empirical mean.
% They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model,
% as the statistically best possible arm (\emph{i.e.}, the highest upper confidence bound) is selected.
% % the arm with the highest upper confidence bound is expected to be the true best arm, and is played.

% More formally, for one end-user, let $N_k(t) = \sum_{\tau=1}^t \mathbbm{1}(C(\tau) = k)$ be the number of times channel $k$ was selected up-to time $t \geq 1$.
% The empirical mean estimator of channel $k$ is defined as the mean reward obtained by selecting it up to time $t$, $\widehat{\mu_k}(t) = \left(1 / N_k(t)\right) \sum_{\tau=1}^t r_k(\tau) \mathbbm{1}(C(\tau) = k) $.
% For \UCB, the \emph{confidence} term is $B_k(t) = \sqrt{\alpha \log(t) / N_k(t)}$,
% giving the upper confidence bound $U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, which is used by the end-user to decide the channel for communicating at time step $t+1$: $C(t+1) = \arg\max_{1\leq k \leq K} U_k(t)$.
% \UCB{} is called an \emph{index policy}.

The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer02}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{Bubeck12}.
% This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
%
% \todo[inline]{Maybe in the implementation of UCB the $t$ $\log(t)$ should be the total amount of time the user tried to communicate? It is already the case, but we should say so.}
In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.

% \paragraph{Thompson Sampling}

% Thompson Sampling \cite{Thompson33} was introduced early on, in $1933$ as the very first bandit algorithm, in the context of clinical trials (in which each arm models the efficacy of one treatment across patients). Given a prior distribution on the mean of each arm, the algorithm selects the next arm to draw based on samples from the \emph{conjugated} posterior distribution, which for Bernoulli rewards is a Beta distribution.

% A Beta prior $\mathrm{Beta}(a_k(0)=1,b_k(0)=1)$ (initially uniform) is assumed on $\mu_k \in [0, 1]$, and at time $t$ the posterior is $\mathrm{Beta}(a_k(t),b_k(t))$.
% After every channel selection, the posterior is updated to have $a_k(t)$ and $b_k(t)$ counting the number of successful and failed transmissions made on channel $k$.
% So if the \emph{ACK} message is received, $a_k(t+1) = a_k(t) + 1$, and $b_k(t+1) = b_k(t)$, otherwise $a_k(t+1) = a_k(t)$, and $b_k(t+1) = b_k(t) + 1$.
% Then, the decision is done by \emph{sampling} an \emph{index} for each arm, at each time step $t$, from the arm posteriors: $X_k(t) \sim \mathrm{Beta}(a_k(t), b_k(t))$, and the chosen channel is simply the channel $C(t+1)$ with highest index $X_k(t)$. For this reason, Thompson Sampling is called a \emph{randomized index policy}.

The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12Thompson}.
It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov16,Mitton16}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.

@@

\paragraph{\emph{Adversarial} bandit algorithms?}
%
Instead of using MAB algorithms assuming a stochastic hypothesis on the system, we could try to use MAB algorithms designed to tackle a more general problem, that makes no hypothesis on the interfering traffic.
The \emph{adversarial MAB} algorithms is a broader family, and a well-known and efficient example is the $\mathrm{Exp}3$ algorithm \cite{Bubeck12}.
Empirically, the $\mathrm{Exp}3$ algorithm turned out to perform worse than both \UCB{} and TS in the same experiments.
%(see Section \ref{sub:41:numericalResults}).
Contrarily to the two stochastic algorithms, the use of $\mathrm{Exp}3$ is correctly justified, even in the non-stationary and non-\iid, as its performance guarantee are true in \emph{any} setting.
But it is not so surprising that it performs worse, as the theoretical performance guarantees of adversarial MAB algorithms are an order of magnitude
%\footnote{We preferred not to talk about \emph{regret} in this study, but in a few words: it is a measure of how \emph{bad} the algorithm performed in terms of its accumulated rewards, in comparison to the best possible policy, and should be as small as possible. For stochastic algorithms, being ``efficient'' means having a regret bounded as $R_T = \mathcal{O}(\log T)$, but for adversarial algorithms, it means having $R_T = \mathcal{O}(\sqrt{K T})$.}
worse than the one for stochastic ones.
% (in their respective case of application).
More is left on this aspect for our future work.

@@

\begin{figure}[!t]
    \centering
    \subfloat[10\% of smart devices]{\includegraphics[scale=0.5]{10intelligent.eps}
    \label{fig:41:10intelligent}}
    \hfill
    \subfloat[30\% of smart devices]{\includegraphics[scale=0.5]{30intelligent.eps}
    \label{fig:41:30intelligent}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of smart devices]{\includegraphics[scale=0.5]{50intelligent.eps}
    \label{fig:41:50intelligent}}
    \hfill
    \subfloat[100\% of smart devices]{\includegraphics[scale=0.5]{100intelligent.eps}
    \label{fig:41:100intelligent}}
    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme references without learning or oracle knowledge, when the proportion of smart end-devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.65]{perf_learning.eps}
    \caption{Learning with \UCB{} and Thomson Sampling, with many smart devices.}
    \label{fig:41:perf_learning}
\end{figure}

@@

We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static device, to check that learning does not help in this case, but does not decrease performance either.
% And we should include another example with a different scenarios, maybe $(0.3, 0.3, 0.3, 0.015, 0.015, 0.014, 0.014, 0.014, 0.014, 0.014)$ (R\'emi already did the simulations).

\begin{figure}[!t]
    \centering
    \subfloat[10\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_10.eps}
    \label{fig:41:ch2_10}}
    \hfill
    \subfloat[30\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_30.eps}
    \label{fig:41:ch2_30}}
    \vskip\baselineskip
    \vspace*{-20pt}
    \subfloat[50\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_50.eps}
    \label{fig:41:ch2_50}}
    \hfill
    \subfloat[100\% of intelligent devices]{\includegraphics[scale=0.5]{ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption{Performance of the UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$ (limit scenario).}
    \label{fig:41:figure4appendix}
\end{figure}
